Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-02 fib2.tex
more edits

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Our simulation is in effect that of a $(2+1)$-dimensional topological quantum field theory, which allows for topological processes such as pair-creation, braiding, and fusion of anyonic charges. This is modelled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside.  For more details of an analogous model, see~\cite{Brell2013}.
\dude{something about observables vs. states.}

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009},
and follows a similar strategy to the hard-decision renormalization group (RG) decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-decode.pdf}
\caption{The decoder works by maintaining a set of disjoint clusters as rooted trees.
(a) At the initial clustering stage, these trees are formed 
from neighboring sites that contain charges. Within each cluster, the 
charges are transported to the root of the tree (chosen 
arbitrarily), and their combined charge measured. The direction of transport 
(towards the root) is denoted by arrows.
(b) At each successive round, all trees are grown in 
every direction, and overlapping trees are joined. Again, any charges 
within a cluster are transported to the root of the 
tree and measured. Any clusters with vacuum total charge are deleted.
\label{f:decode}
\end{center}
\end{figure}

@@

In FIG. \ref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}
First, (a) we show the state as a tensor product of two pairs $\tau\tau\otimes\tau\tau$.
Then, (b) these two pairs are embedded into a single larger fusion
tree $\tau\tau\tau\tau$.
This is continuosly deformable into the picture in (c).
It is now clear the two braid operations required so
that the observables are represented on the fusion tree.
The general algorithm and datastructures used
for this proceedure will be discussed elsewhere.

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, applying these methods to more realistic models such as concrete microscopic spin models or models with non-topological features would give direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. 
Another potential direction is to incorporate tensor network methods for simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
Finally, perhaps the most important open question is whether \emph{fault-tolerant} decoders can be designed for nonabelian anyon codes.

>>>>

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Our simulation is in effect that of a $(2+1)$-dimensional topological quantum field theory, which allows for topological processes such as pair-creation, braiding, and fusion of anyonic charges. This is modelled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside.  For more details of an analogous model, see~\cite{Brell2013}.

As was noted in~\cite{Pfeifer2014}, when dealing with two-dimensional 
anyon dynamics, it is not sufficient to consider only
the algebraic (fusion/braiding) properties of the anyons,
one must also account for the algebra of observables.

@@

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

\begin{figure}[t!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-decode.pdf}
\caption{The decoder works by maintaining a set of disjoint clusters as rooted trees.
(a) At the initial clustering stage, these trees are formed 
from neighboring sites that contain charges. Within each cluster, the 
charges are transported to the root of the tree (chosen 
arbitrarily), and their combined charge measured. The direction of transport 
(towards the root) is denoted by arrows.
(b) At each successive round, all trees are grown in 
every direction, and overlapping trees are joined. Again, any charges 
within a cluster are transported to the root of the 
tree and measured. All clusters with vacuum total charge are deleted.
\label{f:decode}
}
\end{center}
\end{figure}

@@

In \Fref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}
First, (a) we show the state as a tensor product of two pairs $\tau\tau\otimes\tau\tau$.
Then, (b) these two pairs are embedded into a single larger fusion
tree $\tau\tau\tau\tau$.
This is continuosly deformable into the picture in (c).
It is now clear the two braid operations required so
that the observables are represented on the fusion tree.
The general algorithm and datastructures used
for this proceedure will be discussed elsewhere.

@@

There are several interesting avenues for further research. 
Although it is not completely obvious how to do so, applying these methods to more realistic models such as concrete microscopic spin models or models with non-topological features would give direct insight into practical error rates needed for topological quantum memories in non-Abelian systems. 
\dude{This doesn't make much sense to me:} Another potential direction is to incorporate tensor network methods for simulating anyons~\cite{Pfeifer2012, Singh2014} to improve the speed and scale of the simulations. 
Finally, perhaps the most important open question is whether 
\emph{fault-tolerant} decoders can be designed for nonabelian anyon codes. 
\dude{Another important question is wether this error
correction process can be mapped to a classical stat mech problem, like RBIM.}
